**Department Invited Speakers Do Not Reflect Trainee Diversity**

### 2019-09-05 submission/manuscript.Rmd
work on results & discussion

@@

**Data collection and demographic determination.** Historical information on invited speakers and their hosts were obtained from the departmental records of five academic years (Fall 2014 - Spring 2019). Each speaker was only counted once and any departmental faculty members or those listed as a "host" at any point could not also be considered "invited speakers". The list of current trainees (Spring 2019), was obtained from department listservs and included masters students, doctoral students, and post-doctoral fellows.
  Demographics for invited speakers, hosts, and trainees were hand-coded using personal knowledge, photos, and CVs. Presenting gender was assigned using a binary system (man/woman). Race/ethnicity was assigned using the current U.S. Census definition where "Caucasian" includes those of Middle Eastern, European, and Russian descent. Historically Underrepresented Minority (HURM) individuals are restricted to those with African-American, Indigenous and/or Hispanic heritage. International indviduals were either visiting the US at the time of their talk, or immigrated to the US as an adult. Individual assignments can be found in the anonymized dataset.

@@

+ Gender
    + decreasing representation of women in each role (as prestige increases)
    + compare to global stats of Micro resources 60:40 (Elseiver)
    + previous study at 50 US universities focusing on genders
    + discuss factors that decrease speaking invitations to women

+ Under represented minorites
    + overrepresentation of Caucasian individuals as Faculty & Speakers, relative to trainees
    + decline in HURM & International as Fac/Speak, relative to trainees
    + inclination to lump HURM & International together to improve diversity stats
    + important to distinguish separately as both face different issues in the US & academy

>>>>

+ introduce invited speaker series - what are the goals, who attends, etc.

@@

**Data collection and demographic determination.** Historical information on invited speakers and their hosts were obtained from the departmental records of five academic years (Fall 2014 - Spring 2019). Each speaker was only counted once and any departmental faculty members or those listed as a "host" at any point could not also be considered "invited speakers". The list of faculty hosts was used as a proxy for faculty demographics since each faculty member has the opportunity to invite one speaker per year. The list of trainees (Spring 2019), was obtained from department listservs and included masters students, doctoral students, and post-doctoral fellows.
  Demographics for invited speakers, hosts, and trainees were hand-coded using personal knowledge, photos, and CVs. Presenting gender was assigned using a binary system (man/woman). Race/ethnicity was assigned using the current U.S. Census definition where "Caucasian" includes those of Middle Eastern, European, and Russian descent. Historically Underrepresented Minority (HURM) individuals are restricted to those with African-American, Indigenous and/or Hispanic heritage. International indviduals were either visiting the US at the time of their talk, or immigrated to the US as an adult. Individual assignments can be found in the anonymized dataset.

@@

To understand the representation of women, we compared the proportion of individuals presenting as women in each academic role. At the trainee level, more than half of students and postdoctoral fellows were women. That dropped to X% of faculty hosts and X% of the invited speakers. [what about lectureships?] The proportion of women as faculty hosts and speakers is equvialent to global estimates of women microbiologists (Elseiver). Several papers have investigated the representation of women at scientific conferences, however, we only identified one that focused on invited speakers at university-sponsored events (Nittrouer, 2018). [Compare to our findings]

Assumptions about competency and dedication may negatively impact the liklihood of women microbiologists being invited as speakers. The  dedication of women with family to their work is also more likely to be questioned than that of their colleagues, even men who also have children. For instance, the percieved prioritization and commitments of women to family over work may cause faculty to doubt their acceptance of a speaking invitation (despite the prestigous nature of these inviations), causing the faculty member to invite a different colleague who they feel is less likely to turn them down. Departments have different processes and criteria for selecting invited speakers, but want to bring the best scientists possible. It may be that the definition of "best" poses a problem to women, who need three-times as many publications as their men collegues to be considered equally competent. Some departments only invite tenured faculty, which severely limits the number of potential women speakers considerably. Alternately, pre-tenure faculty members invite presitgous, tenured faculty in their field to network and secure letters for their own tenure package. The increased burden of women to prove competency decreases their likihood to be considered for either tenure and as possible source of tenure letters.

Our analysis identified an overrepresentation of Caucasian individuals as hosting faculty and invited speakers, relative to the proportion of Caucasian trainees. Diversity definitions vary according to the goals and population in question. However, in the United States, there is an inclination to lump individuals of HURM and international backgrounds together to improve diversity statistics. We argue that it is important to distinguish these two groups separately since each face different issues in the US and the academy. For instance, international scientists face more issues with visas while HURMs must contend with a country who systematically shuts them out despite an infrastructure that was built on their historical land and labor. For this reason, and to provide more resolution to the analysis, individuals were also given additional designations such as HURM and international. We observed declines in the representation of HURM and international faculty and speakers relative to the trainees. [Lectureships].
